Computational tools for representation theory of affine Lie algebras

### 2009-07-09 AffineLieAlgebras/ACMS2009/notes.tex
Continue work on WZW models section

...
\section{CFT}
\label{sec:cft}

@@

For example, following fields are Weyl-invariant:
\begin{equation}
  \label{eq:68}
  O_a=\int \Phi_{\Delta}e^{2a\phi}\,d^2x,\quad \mbox{where}\;\Delta+a(Q-a)=1
\end{equation}
$O_a$ - is the main observable of string theory with gravitational dimension $\delta_a=-\frac{a}{b}$.
\begin{equation}
  \label{eq:69}
  \int\langle O_{a_1}\dots O_{a_n}\rangle_{\mu} e^{-\mu A}d\mu=\langle O_{a_1}\dots O_{a_n}\rangle_{A}=A^{\sum \delta_k},\quad \delta_k=\frac{a_k}{b}
\end{equation}

\subsection{Conclusion}
\label{sec:conclusion}

@@

\section{Conformal Field Theory}
\label{sec:conf-field-theory}

>>>>

...
\section{Conformal Field Theory}
\label{sec:cft}

\subsection{General properties}
\label{sec:general-properties}

@@

For example, following fields are Weyl-invariant:
\begin{equation}
  \label{eq:68}
  O_a=\int \Phi_{\Delta}e^{2a\phi}\,d^2x,\quad \mbox{where}\;\Delta+a(Q-a)=1
\end{equation}
$O_a$ - is the main observable of string theory with gravitational dimension $\delta_a=-\frac{a}{b}$.
\begin{equation}
  \label{eq:69}
  \int\langle O_{a_1}\dots O_{a_n}\rangle_{\mu} e^{-\mu A}d\mu=\langle O_{a_1}\dots O_{a_n}\rangle_{A}=A^{\sum \delta_k},\quad \delta_k=\frac{a_k}{b}
\end{equation}
\subsection{Connection with string theory}

@@

\subsection{Wess-Zumino-Novikov-Witten models}
\label{sec:WZNW}

Let's start with non-linear $\sigma$-model.
\begin{equation}
  \label{eq:48}
  S_0=\frac{1}{4a^2}\int d^2x\; Tr' (\partial^{\mu}g^{-1}\partial_{\mu}g)
\end{equation}
Here $a^2>0$ is a positive parameter, $g(x)\in G$ - field on Lie group $G$, which we will assume to be semi-simple and by $Tr'$ we have denoted representation independent trace of Lie algebra $\mathfrak{g} $
\begin{equation}
  \label{eq:70}
  Tr'=\frac{1}{x_{rep}}Tr
\end{equation}
$x_{rep}$ is the Dynkin index of the representation.

In this model the conformal invariance is lost at the quantum level, as it can be seen along the lines of previous sections.
!!! Show it by example !!!

Also we can see that holomorphic and antiholomorphic currents aren't conserved individually. The equations of motion are
\begin{equation}
  \label{eq:71}
  \partial^{\mu}(g^{-1}\partial_{\mu}g)=0
\end{equation}
The currents are
\begin{equation}
  \label{eq:currents}
  J_{\mu}=g^{-1}\partial_{\mu}g
\end{equation}
or in the complex coordinates
\begin{equation}
  \label{eq:74}
  \begin{matrix}
    & \tilde{J}_z=g^{-1}\partial_z g, & \tilde{J}_{\bar{z}}=g^{-1}\partial_{\bar{z}}g\\
    & \partial_z \tilde{J}_{\bar{z}}+\partial_{\bar{z}}\tilde{J}_z=0 & \\
  \end{matrix}
\end{equation}
The terms of the equation of motion can not be null separately because $\partial_{\mu}(\epsilon^{\mu\nu}J_{\nu})\neq 0$.

So we add WZ-term to the action and redefine the currents in more appropriate way
\begin{equation}
  \label{eq:72}
  J_z=\partial_z g\;g^{-1}, \qquad J_{\bar{z}}=g^{-1}\partial{\bar z}g
\end{equation}

Wess-Zumino term has the following form
\begin{equation}
  \label{eq:73}
\Gamma=  - \frac{i }{24\pi} \int_{B}\epsilon_{ijk} Tr'\left(
    \tilde g^{-1}\frac{\partial \tilde g}{\partial y^i}
      \tilde g^{-1}\frac{\partial \tilde g}{\partial y^j}
      \tilde g^{-1}\frac{\partial \tilde g}{\partial y^k}\right) d^3y
\end{equation}

This term is defined on the three-dimensional manifold $B$ whose boundary is original two-dimensional space. By $\tilde{g}$ we have denoted the extension of the field $g$ to $B$. This extension isn't unique. In a compactified three-dimensional space a compact two-dimensional manifold delimits two three-dimensional manifolds. The difference between this two choices $\Delta\Gamma$ is given by the right hand side of (\ref{eq:73}) with the integration range extended over the whole compact three-dimensional space. Since the latter is topologically equivalent to the three-sphere we have
\begin{equation}
  \label{eq:75}
\Delta\Gamma=  - \frac{i }{24\pi} \int_{S^3}\epsilon_{ijk} Tr'\left(
    \tilde g^{-1}\frac{\partial \tilde g}{\partial y^i}
      \tilde g^{-1}\frac{\partial \tilde g}{\partial y^j}
      \tilde g^{-1}\frac{\partial \tilde g}{\partial y^k}\right) d^3y
\end{equation}
$\Delta\Gamma$ is defined modulo $2\pi i$, so the Euclidean functional integral with weight $exp(-\Gamma)$ is well-defined. Clearly, any coupling constant multiplying this term must be an integer.

We then consider the action
\begin{equation}
  \label{eq:76}
  S=S_0+k\Gamma
\end{equation}
where $k$ is an integer. 
The equation of motion for the full action (\ref{eq:76}) is
\begin{equation}
  \label{eq:77}
  \partial^{\mu}(g^{-1}\partial_{\mu}g)+\frac{a^2 ik}{4\pi}\epsilon_{\mu\nu}\partial^{\mu}(g^{-1}\partial^{\nu}g)=0
\end{equation}
It can be rewritten using complex coordinates as
\begin{equation}
  \label{eq:78}
  (1+\frac{a^2 k}{4\pi})\partial_z(g^{-1}\partial_{\bar z}g)+(1-\frac{a^2 k}{4\pi})\partial_{\bar z}(g^{-1}\partial_z g)=0
\end{equation}
We see that when $a^2=\frac{4\pi}{k}$ we have the desired conservation law
\begin{equation}
  \label{eq:79}
  \partial_z(g^{-1}\partial{\bar z}g)=0
\end{equation}
For currents
\begin{equation}
  \label{eq:4}
  \partial_{\bar z}J=0,\quad \partial_z \bar J=0
\end{equation}

The solution of classical equation is
\begin{equation}
  \label{eq:80}
  g(z,\bar z)=f(z)\bar f(\bar z)
\end{equation}
for arbitrary functions $f(z)$ and $\bar f (\bar z)$.

The separate conservation of the currents $J_z,\; J_{\bar z}$ implies the invariance of the action under the transformations
\begin{equation}
  \label{eq:81}
   g(z,\bar z)\to \Omega(z)g(z,\bar z)\bar \Omega^{-1}(\bar z)
\end{equation}
where $\Omega,\;\bar \Omega \in G$. So we have local $G(z)\times G(\bar z)$ invariance.

In order to move analysis to quantum level we rescale the currents
\begin{equation}
  \label{eq:82}
  J(z)\equiv -k \partial_zg g^{-1}\quad \bar J(\bar z)=k g^{-1}\partial_{\bar z}g
\end{equation}
So the variation of the action under infinitesimal transformation $\Omega=1+\omega,\; \bar \Omega =1+\bar \omega$ is
\begin{equation}
  \label{eq:83}
  \delta_{\omega,\bar\omega}S=\frac{i}{4\pi}\oint dz Tr' (\omega(z)J(z))-\frac{i}{4\pi}\oint d\bar z Tr'(\bar\omega(\bar z)\bar J(\bar z))
\end{equation}
Expanding the currents as
\begin{equation}
  \label{eq:85}
  \begin{aligned}
    J=\sum J^a t^a,\bar J=\sum \bar J^a t^a \\
    \omega=\sum \omega^a t^a\\
  \end{aligned}
\end{equation}
we get
\begin{equation}
  \label{eq:86}
  \delta_{\omega,\bar \omega}S=-\frac{1}{2\pi i}\oint dz \sum\omega^a J^a+\frac{1}{2\pi i} \oint d\bar z \sum \bar \omega^a \bar J^a
\end{equation}
We can also obtain Ward identities $\delta\left< X\right>=\left<(\delta S)X\right>$
\begin{equation}
  \label{eq:87}
  \delta_{\omega,\bar \omega}\left< X \right>=-\frac{1}{2\pi i}\oint dz \sum\omega^a \left< J^a X\right>+
  \frac{1}{2\pi i} \oint d\bar z \sum \bar \omega^a \left< \bar J^a X\right>
\end{equation}
Then for the currents we have
\begin{equation}
  \label{eq:88}
  \delta_{\omega}J=[\omega,J]-k\partial_z\omega,\quad \delta_{\omega}J^a=\sum i f_{abc}\omega^b J^c-k\partial_z\omega^a
\end{equation}
The operator product expansion for currents has the form
\begin{equation}
  \label{eq:89}
  J^a(z) J^b(w) \sim \frac{k\delta_{ab}}{(z-w)^2}+\sum i f_{abc}\frac{J^c(w)}{(z-w)}
\end{equation}
Expanding currents into series we get
\begin{equation}
  \label{eq:90}
  \begin{aligned}
    J^a(z)=\sum_{n\in \mathbb Z}z^{n-1}J^a_n\\    
    \left[J^a_n,J^b_m\right]=\sum_c i f^{abc}J^c_{n+m}+kn\delta^{ab}\delta_{n+m,0}  
  \end{aligned}
\end{equation}
We see that components of currents realize affine Lie algebra $\hat g$.

Energy-momentum tensor is introduced with Sugawara construction as the sum of normal-ordered products of current components
\begin{equation}
  \label{eq:6}
  T(z)=\frac{1}{2(k+h^v)}\sum_a N(J^a J^a)(z)
\end{equation}
Here $h^v$ is the dual Coxeter number.

Energy-momentum tensor can be expanded into modes $L_n$
\begin{equation}
  \label{eq:91}
  L_n=\frac{1}{2(k+h^v)}\sum_a\sum_m:J^a_m J^a_{n-m}:
\end{equation}
The commutation relations for the modes $L_n$ are
\begin{equation}
  \label{eq:92}
  \begin{aligned}
    \left[L_n,L_m\right]=(n-m)L_{n+m}+\frac{c}{12}(n^3-n)\delta_{n+m,0}\\
    \left[L_n,J^a_m\right]=-mJ^a_{n+m}
  \end{aligned}
\end{equation}

So the Sugawara construction is the way to embed Virasoro algebra into the universal enveloping of affine Lie algebra.

Complete chiral algebra of WZW-model is equal to semidirect product $Vir\ltimes \hat g$
